Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-12-21 apj.tex
new version for limited viewing

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The two color parameters exhibit the behavior expected from the two-parameter
\citet{1999PASP..111...63F} dust model.
The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
the two parameters
plus a detection of a third parameter at
%--- table.py
$>95$\% confidence;
%-----
this new parameter is responsible for supernova color diversity
beyond those tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

Our analysis sample is comprised by the
168
supernovae that have the data coverage to 
give photometric and spectroscopic statistics described above.
The data are given in Table~\ref{data:tab} \textcolor{red}{[TABLE CAN GO IF THERE IS A NICO PAPER]}.
The peak magnitude uncertainties do have covariance, which is accounted
for in the analysis; only the standard deviation is included in the table.

@@

\color{red}
For $N$ supernovae there are $8N$ observables.  For the top-level model parameters, there are $3N$ spectral parameters, $2(N-2)$
color parameters, $(N-1)$ magnitude offsets,  $5 \times 6$ global coefficients, and $10$ parameters that describe the intrinsic
$C_c$.  For $N=168$ supernovae, there are 1344 observables and 1043 top-level parameters.
\color{black}

@@

\subsection{Results}
\label{results:sec}
The posterior of the model parameters is evaluated using Hamiltonian Monte Carlo as implemented in
STAN \citep{stan}.  We run four chains, each with 10000 iterations of which
half are used for warmup.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$ \citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{eff}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{eff} \gg 100$ for all parameters, indicating that are all densely sampled.
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig}, \ref{global2:fig}, \ref{global3:fig}, \ref{global4:fig},
\ref{global5:fig}.  We find no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
We have rurun the analysis with a variety of initial conditions, including one with all the $\gamma$'s equal to zero, except for a small positive 
$\gamma^0_0$ and small negative $\gamma^1_0$ to satisfy our limit conditions.  Parameter credible intervals
remain consistent within the 68\% credible intervals.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/coeff0.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $U$ band.
            The contours shown here and in future plots represent 1-$\sigma$ in the parameter distribution (i.e.\ they should be
            projected onto the corresponding 1-d parameter axis), not to 68\%, 95\%, etc.\
            enclosed probability.  \label{global1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/coeff1.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $B$ band.
 \label{global2:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/coeff2.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $V$ band.
 \label{global3:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.5in]{output11/coeff3.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $R$ band.
 \label{global4:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
         \includegraphics[width=5.5in]{output11/coeff4.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $I$ band.
 \label{global5:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/residual.pdf} 
            \caption{Predicted versus observed colors.  Overplotted is a line with slope equal to one.
            \label{residual:fig}}
\end{figure}

@@

The ideogram for the grey offsets, $\Delta$, for all supernovae is shown in Figure~\ref{hist:fig}.  The distribution is non-Gaussian, 
has a standard deviation of
%-----
$0.10$
%-----
mag, and a broad tail in the positive (fainter) direction.
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/Delta_hist.pdf} 
   \caption{Ideogram for the grey offset $\Delta$.  This and other ideograms plotted in this article are normalized to have unit area.
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.  Given that the 
off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
it is not straightforward to present correlation matrix confidences:
the average over Monte Carlo links of the
Cholesky factors will not  generally yield a correlation matrix.  
To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
chain, element by element.
For $UBVRI$ the matrix is
\begin{equation}
C_c(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
\end{array}
 \end{pmatrix} \text{mag}^2.
 \label{mag_cov:eqn}
 \end{equation}
\color{red}
The average correlation matrix is
\begin{equation}
Cor(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
0.12 & -0.39 & -0.22 & 0.27 & 1.00
\end{array}
 \end{pmatrix}.
 \end{equation}
 \color{black}
The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
expressed as the standard deviations and
 correlation matrix
 \begin{equation}
 \sigma(U-V, B-V, V-R, V-I)=
 \begin{pmatrix}
0.068 & 0.034 & 0.023 & 0.053
  \end{pmatrix} \text{mag}
 \label{color_sd:eqn}
   \end{equation}
 \begin{equation}
 Cor(U-V, B-V, V-R, V-I)=
\begin{pmatrix}
\begin{array}{rrrr}
1.000 & 0.602 & -0.353 & -0.317 \\
0.602 & 1.000 & -0.197 & 0.082 \\
-0.353 & -0.197 & 1.000 & 0.647 \\
-0.317 & 0.082 & 0.647 & 1.000
\end{array}
\end{pmatrix}.
  \label{color_cor:eqn}
 \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N, 2007ApJ...659..122J}, for cases where direct comparison can be made.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output11/perobject_corner.pdf} 
   \caption{Distributions for the supernova parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^1}(B-V)$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^0}(B-V)$ are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V)) =\\
\begin{pmatrix}
\begin{array}{rrrrrr}
1.00 & 0.00 & -0.06 & -0.05 & 0.11 & 0.01 \\
0.00 & 1.00 & 0.10 & -0.26 & -0.07 & -0.01 \\
-0.06 & 0.10 & 1.00 & -0.14 & -0.17 & -0.03 \\
-0.05 & -0.26 & -0.14 & 1.00 & 0.03 & 0.03 \\
0.11 & -0.07 & -0.17 & 0.03 & 1.00 & 0.05 \\
0.01 & -0.01 & -0.03 & 0.03 & 0.05 & 1.00
\end{array}
\end{pmatrix}.
\end{multline}
We find weak correlations between our color-excess parameters and the input features, the strongest being
%---
$-0.17$
%---
 between
$E_\gamma(B-V)$ and $EW_{Si}$.
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\subsection{The Two Extrinsic Parameters $k_0$ and $k_1$}
The two sets of  $\gamma^i$ parameters  are significantly non-zero. 
None of the 20000 links of 
our Monte Carlo chains for $\gamma$ exceed 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
influence of $k_0$ and $k_1$  on supernova magnitudes
with probability $(1-5\times 10^{-5})$.

\color{blue}
Recall that there are degeneracies between the intrinsic (Eqn.~\ref{ewsiv:eqn}) and
extrinsic (Eqn.~\ref{dust:eqn}) parameters of our model.  Part of that degeneracy is
broken by imposing $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$,
but this choice does not specify the $k_0^0$, $k_1^0$ corresponding to supernovae experiencing no extrinsic effects.  The true extrinsic
extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
Our model as analyzed cannot directly determine the above extrinsic properties per supernova.

Nevertheless, we are able to examine aggregate properties of the extinction.  The extrinsic extinction and color relation can be
written as
$$A_X(k_0,k_1) = R^T_V \left( A_B(k_0,k_1) -  A_V(k_0,k_1)\right) -R^T_V \left(A_B(k^0_0,k^0_1) + A_V(k^0_0,k^0_1)\right)+  A_X(k^0_0,k^0_1) .
$$
Making the assumption that $R^T_V$ is the same for all supernovae (shown later to be approximately true for dust laws),
it can be determined as the slope of the relation between $A_X$ and $E(B-V)$.
\color{black}
Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of
%----- fitz11.py
$R_V=2.77 \pm 0.08$.
%----
\color{blue}
The slope uncertainties are the $68$\% credible interval from the Bayesian analysis and does not incorporated goodness
 of fit.  Indeed,
the full set of supernovae does not coincide with the expectations of a single linear law; the supernova population rather appears to be peaked
on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
but can have values that extend toward lower values.

%\begin{comment}
%\color{blue}
%Our model alone cannot determine $R^T_V = A^T_V/E^T(B-V)$ for individual objects, as $k_0^0$ and $k_1^0$ are unconstrained.
%For a qualitative view, we choose zeros  such that the $V$-band extinction and color-excess of the bluest supernova roughly correspond to $E(B-V)=0$,
%$A_V=0$, and such that the ridge line 
%has close to constant $R^T_V$.
%\color{black}
%With this ansatz, we
%estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
% \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
%With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
%We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
%of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
%A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
%for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
%\color{black}
%
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/avebv.pdf}
   %\includegraphics[width=2.8in]{output11/rv.pdf}
   \caption{
	\color{red}
   $A_V$ versus $E(B-V)$, and the slope $R_V$ of the linear fit to the data.
   %Right: Shown for qualitative purposes,
   %$R^T_V=A^T_V/E^T(B-V)$ 
   %versus $E^T(B-V)$ for the supernova sample, given an ansatz for the definition of an intrinsic supernova. 
   %Overplotted is the mean and standard deviation of those points in equally divided bins.
   Overplotted the expected values for the \citetalias{1999PASP..111...63F} model
   for specific choices of $R^F$.  Note that these lines are not straight on the left plot.
   \color{black}
   \label{avebv:fig}}
\end{figure}
%\end{comment}

We now compare our model and its color-parameters $k_0$ and $k_1$ with the expectations of  dust laws,
in contrast to  empirical properties derived from our model.
A direct analysis of the  \citetalias{1999PASP..111...63F} dust model is presented later
in \S\ref{model0:sec}.

@@

The \citet{1989ApJ...345..245C} dust law can be expressed as a two-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V).
\label{f99:eqn}
\end{equation}
The \citetalias{1999PASP..111...63F} model is approximately described by a linear model with
the above form:
\color{red}
for the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak as seen through our synthetic passbands,
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$ 
\color{black}
Over the ranges of
\color{red}
 $R^F$ and $A^F_V$,
\color{black}
and wavelengths under consideration,  the values of the components  vary by $<5$\%.

>>>>

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Extending the model to three color parameters gives almost identical results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter is responsible for supernova color diversity
beyond those tracked by the spectral features and two-parameter dust models.
\end{abstract}

@@

Our analysis sample is comprised by the
172
supernovae that have the data coverage to 
give photometric and spectroscopic statistics described above.
The data are given in Table~\ref{data:tab} \textcolor{red}{[TABLE CAN GO IF THERE IS A NICO PAPER]}.
The peak magnitude uncertainties do have covariance, which is accounted
for in the analysis; only the standard deviation is included in the table.

@@

\color{red}
For $N$ supernovae there are $8N$ observables.  For the top-level model parameters, there are $3N$ spectral parameters, $2(N-2)$
color parameters, $(N-1)$ magnitude offsets,  $5 \times 6$ global coefficients, and $10$ parameters that describe the intrinsic
$C_c$.  For $N=172$ supernovae, there are 1376 observables and 1067 top-level parameters.
\color{black}

@@

\subsection{Results}
\label{results:sec}
The posterior of the model parameters is evaluated using Hamiltonian Monte Carlo as implemented in
STAN \citep{stan}.  We run eight chains, each with 5000 iterations of which
half are used for warmup.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$ \citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{eff}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{eff} \gg 100$ for all parameters, indicating that are all densely sampled.
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig} -- \ref{global5:fig}.  We find no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
We have rurun the analysis with a variety of initial conditions, including one with all the $\gamma$'s equal to zero, except for a small positive 
$\gamma^0_0$ and small negative $\gamma^1_0$; the 68\% credible intervals
remain consistent.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/coeff0.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $U$ band.
            The contours shown here and in future plots represent 1-$\sigma$ in the parameter distribution (i.e.\ they should be
            projected onto the corresponding 1-d parameter axis), not to 68\%, 95\%, etc.\
            enclosed probability.  \label{global1:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/coeff1.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $B$ band.
 \label{global2:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/coeff2.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $V$ band.
 \label{global3:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
      \includegraphics[width=5.2in]{output11/coeff3.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $R$ band.
 \label{global4:fig}}
\end{figure}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
         \includegraphics[width=5.2in]{output11/coeff4.pdf} 
            \caption{Posterior contours for $c$, $\alpha$, $\beta$, $\eta$, $\gamma^0$, $\gamma^1$, and $\sigma$ in the $I$ band.
 \label{global5:fig}}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/residual.pdf} 
            \caption{Predicted versus observed colors.  Overplotted is a line with slope equal to one.
            \label{residual:fig}}
\end{figure}

@@

The ideogram for the grey offsets, $\Delta$, for all supernovae is shown in Figure~\ref{hist:fig}.  The distribution is non-Gaussian, 
has a standard deviation of
%-----
$0.10$
%-----
mag, and a broad tail in the positive (fainter) direction.
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/Delta_hist.pdf} 
   \caption{Ideogram for the grey offset $\Delta$.  This and other ideograms plotted in this article are normalized to have unit area.
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude dispersions are captured in $C_c$.  The diagonal elements are captured by the $\sigma$ parameters;
the residual intrinsic dispersion ranges from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.  Given that the 
off-diagonal elements of $C_c$ are parameterized by the Cholesky factors of a correlation matrix rather than the matrix elements themselves,
it is not straightforward to present correlation matrix confidences:
the average over Monte Carlo links of the
Cholesky factors will not  generally yield a correlation matrix.  
To characterize a typical posterior draw of $C_c$ we use the matrix that is the mean of all covariance realizations in the
chain, element by element.
For $UBVRI$ the matrix is
\begin{equation}
C_c(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
0.0038 & 0.0010 & -0.0002 & -0.0000 & 0.0003 \\
0.0010 & 0.0011 & 0.0002 & -0.0000 & -0.0006 \\
-0.0002 & 0.0002 & 0.0004 & 0.0000 & -0.0002 \\
-0.0000 & -0.0000 & 0.0000 & 0.0002 & 0.0002 \\
0.0003 & -0.0006 & -0.0002 & 0.0002 & 0.0020
\end{array}
 \end{pmatrix} \text{mag}^2.
 \label{mag_cov:eqn}
 \end{equation}
\color{red}
The average correlation matrix is
\begin{equation}
Cor(U,B,V,R,I)=
\begin{pmatrix}
\begin{array}{rrrrr}
1.00 & 0.42 & -0.18 & -0.05 & 0.12 \\
0.42 & 1.00 & 0.26 & -0.09 & -0.39 \\
-0.18 & 0.26 & 1.00 & 0.09 & -0.22 \\
-0.05 & -0.09 & 0.09 & 1.00 & 0.27 \\
0.12 & -0.39 & -0.22 & 0.27 & 1.00
\end{array}
 \end{pmatrix}.
 \end{equation}
 \color{black}
The  covariance of the colors $U-V$, $B-V$, $V-R$, and $V-I$ is
expressed as the standard deviations and
correlation matrix
 \begin{equation}
 \sigma(U-V, B-V, V-R, V-I)=
 \begin{pmatrix}
0.068 & 0.034 & 0.023 & 0.053
  \end{pmatrix} \text{mag},
 \label{color_sd:eqn}
 \end{equation}
 \begin{equation}
 Cor(U-V, B-V, V-R, V-I)=
\begin{pmatrix}
\begin{array}{rrrr}
1.000 & 0.602 & -0.353 & -0.317 \\
0.602 & 1.000 & -0.197 & 0.082 \\
-0.353 & -0.197 & 1.000 & 0.647 \\
-0.317 & 0.082 & 0.647 & 1.000
\end{array}
\end{pmatrix}.
  \label{color_cor:eqn}
 \end{equation}
These color variances are smaller than those derived by \citet{2003A&A...404..901N, 2007ApJ...659..122J}, for cases where direct comparison can be made.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/perobject_corner.pdf} 
   \caption{Distributions for the supernova parameters $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^1}(B-V)$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
   \label{perobject:fig}}
\end{figure}

The Pearson correlation coefficients for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, and $E_{\gamma^0}(B-V)$ are given in the matrix
\begin{multline}
Cor(\Delta, EW_{Ca}, EW_{Si}, \lambda_{Si}, E_{\gamma^0}(B-V), E_{\gamma^1}(B-V)) =\\
\begin{pmatrix}
\begin{array}{rrrrrr}
1.00 & 0.00 & -0.06 & -0.05 & 0.11 & 0.01 \\
0.00 & 1.00 & 0.10 & -0.26 & -0.07 & -0.01 \\
-0.06 & 0.10 & 1.00 & -0.14 & -0.17 & -0.03 \\
-0.05 & -0.26 & -0.14 & 1.00 & 0.03 & 0.03 \\
0.11 & -0.07 & -0.17 & 0.03 & 1.00 & 0.05 \\
0.01 & -0.01 & -0.03 & 0.03 & 0.05 & 1.00
\end{array}
\end{pmatrix}.
\end{multline}
We find weak correlations between our color-excess parameters and the input features, the strongest being
%---
$-0.17$
%---
 between
$E_{\gamma^0}(B-V)$ and $EW_{Si}$.
Recall that $EW_{Si}$ is correlated with light-curve shape, which is correlated with host-galaxy (including dust) properties 
\citep{2000AJ....120.1479H, 2003MNRAS.340.1057S}.
An extensive discussion of the correlations between the spectral features of the SNfactory data set can be found in \citet{chotard:thesis}
and \citet{leget:thesis}.

@@

\subsection{The Two Extrinsic Parameters $k_0$ and $k_1$}
\color{blue}
The two sets of  $\gamma^i$ parameters  are significantly non-zero. 
None of the 20000 links of 
our Monte Carlo chains for $\gamma$ exceed 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
influence of $k_0$ and $k_1$  on supernova magnitudes
with probability $(1-5\times 10^{-5})$.

There is a pair of parameter values,  $k_0^0$ and $k_1^0$, that correspond to supernovae experiencing no extrinsic effect.
The true extrinsic
extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
Our $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$ constraints
do not determine $k_0^0$ and $k_1^0$, and so the extrinsic extinction and color excess per supernova are not products of our analysis.

%our analysis does not specify their values.
%Recall that there are degeneracies between the intrinsic (Eqn.~\ref{ewsiv:eqn}) and
%extrinsic (Eqn.~\ref{dust:eqn}) parameters of our model.  Part of that degeneracy is
%broken by imposing $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$,
%but this choice does not specify the parameter values.

Nevertheless, we are able to examine aggregate properties of the extinction.  The extrinsic extinction and color relation can be
rearranged as
$$A_X(k_0,k_1) = R^T_V \left( A_B(k_0,k_1) -  A_V(k_0,k_1)\right) -R^T_V \left(A_B(k^0_0,k^0_1) + A_V(k^0_0,k^0_1)\right)+  A_X(k^0_0,k^0_1) .
$$
Making the assumption that it is the same for all supernovae ,
 $R^T_V$  can be determined as the slope of the linear relation between $A_X$ and $E(B-V)$.

Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$.
The full set of supernovae does not coincide with the expectations of a single linear law; rather the supernova population is peaked
on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
but can have values that extend toward lower values.
The assumption of a constant $R^T_V$ is not altogether correct; still we perform a linear fit that gives a slope of
%----- fitz11.py
$R^T_V=2.77 \pm 0.08$,
%----
where the uncertainty is the $68$\% credible interval and does not include goodness
 of fit.

%\begin{comment}
%\color{blue}
%Our model alone cannot determine $R^T_V = A^T_V/E^T(B-V)$ for individual objects, as $k_0^0$ and $k_1^0$ are unconstrained.
%For a qualitative view, we choose zeros  such that the $V$-band extinction and color-excess of the bluest supernova roughly correspond to $E(B-V)=0$,
%$A_V=0$, and such that the ridge line 
%has close to constant $R^T_V$.
%\color{black}
%With this ansatz, we
%estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
% \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
%With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
%We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
%of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
%A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
%for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
%\color{black}
%
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/avebv.pdf}
   %\includegraphics[width=2.8in]{output11/rv.pdf}
   \caption{
   $A_V$ versus $E(B-V)$, and the slope $R_V$ of the linear fit to the data.
   %Right: Shown for qualitative purposes,
   %$R^T_V=A^T_V/E^T(B-V)$ 
   %versus $E^T(B-V)$ for the supernova sample, given an ansatz for the definition of an intrinsic supernova. 
   %Overplotted is the mean and standard deviation of those points in equally divided bins.
   Overplotted the expected values for the \citetalias{1999PASP..111...63F} model
   for specific choices of $R^F$. Note that these lines are not straight, which is apparent when comparing to the straight black $R_V=2.78$ line.
   \label{avebv:fig}}
\end{figure}
%\end{comment}

We now compare our model and its color-coefficient parameters $\gamma^0$ and $\gamma^1$ with the expectations of  dust laws.
A direct analysis of the  \citetalias{1999PASP..111...63F} dust model is presented later
in \S\ref{model0:sec}.

@@

The 2-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes dust laws.
For the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the updated
\citet{2007ApJ...663.1187H} SN~Ia template at $B$-band peak, the \citetalias{1999PASP..111...63F} model
gives
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.78\}$ and $b(U,B,V,R,I)=\{  1.64,   1.06,   0.16,  -0.40,  -0.52\}$.
The $F$ superscript is added to distinguish \citetalias{1999PASP..111...63F} parameters.
Over the ranges of
 $R^F$ and $A^F_V$,
and the wavelengths under consideration,  the values of the components  vary by $<5$\%.

Our model can be expressed as
\begin{equation}
A_X = \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  E_{\gamma^0}(B-V) +  \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}  E_{\gamma^1}(B-V).
\end{equation}
Our median-fit values for the $\gamma$ coefficients,
%---- fitz11.py
$\frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V}  =\{4.80 ,   3.89,   2.89,   2.22,   1.59\}$ and
$ \frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V}=\{-3.83 ,  -4.42,  -5.42,  -5.15,  -4.71\}$,
%----
can be transformed to a basis composed of the $a(X)$ and $b(X)$ vectors, and a residual vectors perpendicular to $a$ and $b$.
The result is
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_B-\gamma^0_V} \\
\frac{\gamma^1_X}{\gamma^1_B-\gamma^1_V} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.22 & 2.75 \\
0.81 & -5.50
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the two residual vectors
\begin{align}
\vec{\epsilon}_a &=\{0.14, -0.15, -0.07,  0.04, 0.07\}, \\
\vec{\epsilon}_b & =\{0.12, -0.06, -0.30, 0.30, -0.06\}
\end{align}
contribute  3\% and 4\% respectively to the total  length of the $\gamma$ vectors.
This result indicates that
the allowed variation in $UBVRI$ allowed by the \citetalias{1999PASP..111...63F} model and our best-fit model are closely aligned.

To visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that the four vectors point in different directions, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.
